Cochain level May-Steenrod operations

### 2021-05-25 effective.tex
Rename to May-Steenrod operations

In this section we construct explicit Steenrod-Adem structures on three well know combinatorial $E_\infty$-operads: the Barratt-Eccles operad $\mathcal E$ \cite{berger04combinatorial}, the surjection operad $\mathcal X$ \cite{mcclure03cochain}, and the operad $U(\mathcal M)$ associated to the finitely presented $E_\infty$-PROP $\mathcal M$ \cite{medina2020prop1}. We also define natural and effective Steenrod-Adem structure on the normalized cochains of any simplicial or cubical set, by describing a $U(\mathcal M)$-algebra structures on it.

@@

\begin{figure}
	\begin{tikzcd}
	& & & U(\mathcal M) \arrow[dddr, bend left=10, "\phi^\triangle"'] \arrow[dddrr, bend left=20, "\phi^\square"'] & & \\
	& & \mathcal X \arrow[ru, "SL"] & & & \\
	& \mathcal E \arrow[ur, "TR"] & & & & \\
	\mathcal W \arrow[rrrr, "\psi^\triangle", bend right=10] \arrow[rrrrr, "\psi^\square", bend right=20] \arrow[ur, "\psi_{\mathcal E}"] \arrow[uurr, "\psi_{\mathcal X}"', bend right=35, pos=.6] \arrow[uuurrr, "\psi_{U(\mathcal M)}"', bend right=45, pos=0.6] & & & & \End_{N^\bullet(X)} & \End_{N^\bullet(X)}
	\end{tikzcd}
	\caption{Summary of effective constructions: Steenrod-Adem structures on the Barratt-Eccles $\mathcal E$, surjection $\mathcal X$, and $U(\mathcal M)$ operads, and natural Steenrod-Adem structures on the normalized chains of simplicial and cubical sets. We are abusing notation since the maps $TR$ and $SL$ required different sign conventions.}
	\label{fig: bigsummary}
\end{figure}

\subsection{Barratt-Eccles operad} In this subsection we effectively describe a Steenrod-Adem structure on the Barratt-Eccles operad.

@@

\begin{theorem} \label{thm: Steenrod-Adem on Barratt-Eccles}
	The morphism of $\mathrm{C}$-modules
	\begin{equation*}
	\psi_{\mathcal E} \colon \mathcal W \to \mathcal E
	\end{equation*}
	defines a Steenrod-Adem structure on the Barratt-Eccles operad.
\end{theorem}

@@

In this subsection we introduce a Steenrod-Adem structures on the surjection operad. There are two widely used sign conventions for this operad respectively due to McClure-Smith \cite{mcclure03cochain} and Berger-Fresse \cite{berger04combinatorial}. Using the Berger-Fresse convention, we can define a Steenrod-Adem structure on the surjection operad by composing the map $\psi_{\mathcal E}$ with the \textit{table reduction} quasi-isomorphism $\mathcal E \to \mathcal X$ introduced in Section 1.3 of \cite{berger04combinatorial}. We define an Steenrod-Adem structure on the surjection operad in a convention independent way which recovers the table reduction Steenrod-Adem structure in the Berger-Fresse case.

@@

\begin{theorem} \label{thm: Steenrod-Adem on surjection MS convention}
	The morphism of $\mathrm{C}$-modules
	\begin{equation*}
	\psi_{\mathcal X} \colon \mathcal W \to \mathcal X
	\end{equation*}
	defines a Steenrod-Adem structure on the Surjection operad.
\end{theorem}

@@

We will define a Steenrod-Adem structure $\psi_{U(\mathcal M)} \colon \mathcal W \to U(\mathcal M)$ by composing the Steenrod-Adem structure $\psi_{\mathcal X} : \mathcal W \to \mathcal X$ with a morphism of $\mathrm{C}$-modules ($\mathrm{S}$-modules in fact)
\begin{equation*}
SL \colon \mathcal X \to U(\mathcal M)
\end{equation*}
that we now define. Given a surjection $s \colon \{1, \dots, n+r\} \to \{1, \dots, r\}$ let $SL(s)$ be the element
\begin{center}
	\begin{tikzpicture}[scale=.85]
	\draw (0,0)--(0,-.6) node[below]{$\scriptstyle 1$};
	\draw (0,0)--(.5,.5);
	\draw (-.3, .3)-- (-.2,.5) node[above]{\quad $\scriptstyle 1\ 2\, ...\, k_1$};
	\draw (-.5,.5)--(0,0);
	\node at (.11,.4){$...$};

@@

\begin{theorem} \label{thm: Steenrod-Adem on U(M)}
	The composition $\psi_{U(\mathcal M)}$
	\begin{equation*}
	\begin{tikzcd}[column sep = normal]
	\mathcal W \arrow[r, "\psi_{\mathcal X}"] & \mathcal X \arrow[r, "SL"] & U(\mathcal M)
	\end{tikzcd}
	\end{equation*}
	defines a Steenrod-Adem structure on $U(\mathcal M)$.
\end{theorem}

@@

In this subsection we introduce a natural Steenrod-Adem structure on the normalized cochains of any simplicial set $X$. Since a Steenrod-Adem structure was constructed in the previous section for $U(\mathcal M)$, we only need to describe a natural $U(\mathcal M)$-algebra structure on $N^\bullet(X)$. Using the linear duality functor, it suffices to construct a natural $U(\mathcal M)$-coalgebra structure on $N_\bullet(X)$ which, in turn, can be derived via a Kan extension argument from one on each $N_\bullet(\triangle^n)$. We obtain these coalgebra structures by restricting a full $\mathcal M$-bialgebra structure. An $\mathcal M$-bialgebra structure is specified by three linear maps, the images of the generators
\begin{equation*}
\counit, \quad \coproduct, \quad \product,
\end{equation*}
satisfying the relations in the presentation of $\mathcal M$. For $n \in \mathbb{N}$, define: \vspace*{5pt} \\
(1) The counit $\epsilon \in \Hom(N_\bullet(\triangle^n), \Z)$ by
\begin{equation*}
\epsilon \big( [v_0, \dots, v_q] \big) = \begin{cases} 1 & \text{ if } q = 0, \\ 0 & \text{ if } q>0. \end{cases}
\end{equation*}
(2) The coproduct $\Delta \in \Hom(N_\bullet(\triangle^n), N_\bullet(\triangle^n)^{\otimes2})$ by
\begin{equation*}
\Delta \big( [v_0, \dots, v_q] \big) = \sum_{i=0}^q [v_0, \dots, v_i] \otimes [v_i, \dots, v_q].
\end{equation*}
(3) The product $\ast \in \Hom(N_\bullet(\triangle^n)^{\otimes 2}, N_\bullet(\triangle^n))$ by
\begin{equation*}
\left[v_0, \dots, v_p \right] \ast \left[v_{p+1}, \dots, v_q\right] = \begin{cases} (-1)^{p+|\pi|} \left[v_{\pi(0)}, \dots, v_{\pi(q)}\right] & \text{ if } v_i \neq v_j \text{ for } i \neq j, \\
0 & \text{ if not}, \end{cases}
\end{equation*}
where $\pi$ is the permutation that orders the totally ordered set of vertices, and $(-1)^{|\pi|}$ its sign.

@@

\begin{theorem}
	The commutative diagram
	\begin{equation*}
	\begin{tikzcd}[column sep = small, row sep = small]
	&[18pt] U(\mathcal M) \arrow[dr, "\phi^\triangle", out=0] &[-0pt] \\
	\mathcal W \arrow[ur, "\psi_{U(\mathcal M)}", in=180] \arrow[rr, "\psi^\triangle"] & & \mathrm{End}_{N^\bullet (X)}
	\end{tikzcd}
	\end{equation*}
	defines a natural Steenrod-Adem structure on $N^\bullet(X)$ for any simplicial set $X$.
\end{theorem}

@@

In this subsection we introduce a natural Steenrod-Adem structure on the normalized cochains of any cubical set. It follows the structure of the previous subsection closely. By the same considerations, the desired construction will follow from a natural $\mathcal M$-bialgebra structure on $N_\bullet(\square^n)$. These are determined by three linear maps satisfying the relations in the presentation of $\mathcal M$. For $n \in \mathbb{N}$, define: \vspace*{5pt} \\
(1) The counit $\epsilon \in \Hom(N_\bullet(\square^n), \Z)$ by
\begin{equation*}
\epsilon \left( x_1 \otimes \cdots \otimes x_d \right) = \epsilon(x_1) \cdots \, \epsilon(x_n),
\end{equation*}
where
\begin{equation*}
\epsilon([0]) = \epsilon([1]) = 1, \qquad \epsilon([0, 1]) = 0.
\end{equation*} \vspace*{-6pt} \\
(2) The coproduct $\Delta \in \Hom \left( N_\bullet(\square^n), N_\bullet(\square^n)^{\otimes 2} \right)$ by
\begin{equation*}	
\Delta (x_1 \otimes \cdots \otimes x_n) = 	
\sum \pm \left( x_1^{(1)} \otimes \cdots \otimes x_n^{(1)} \right) \otimes 	
\left( x_1^{(2)} \otimes \cdots \otimes x_n^{(2)} \right),	
\end{equation*}	
where the sign is determined using the Koszul convention, and we are using Sweedler's notation
\begin{equation*}	
\Delta(x_i) = \sum x_i^{(1)} \otimes x_i^{(2)}
\end{equation*}
for the chain map $\Delta \colon N_\bullet(\square^1) \to N_\bullet(\square^1)^{\otimes 2}$ defined by
\begin{equation*}
\Delta([0]) = [0] \otimes [0], \quad \Delta([1]) = [1] \otimes [1], \quad \Delta([0, 1]) = [0] \otimes [0, 1] + [0, 1] \otimes [1].
\end{equation*}
Using that $N_\bullet(\square^n) = N_\bullet(\square^1)^{\otimes n}$, $\Delta$ is the composition
\begin{equation*}
\begin{tikzcd}
N_\bullet(\square^1)^{\otimes n} \arrow[r, "\Delta^{\otimes n}"] & \left( N_\bullet(\square^1)^{\otimes 2}  \right)^{\otimes n} \arrow[r, "sh"] & \left( N_\bullet(\square^1)^{\otimes n} \right)^{\otimes 2}
\end{tikzcd}
\end{equation*}
where $sh$ is the shuffle map that places tensor factors in odd position first. \vspace*{5pt} \\
(3) The product $\ast \in \Hom(N_\bullet(\square^n)^{\otimes 2}, N_\bullet(\square^n))$ by
\begin{align*}
(x_1 \otimes \cdots \otimes x_n) \ast (y_1 \otimes \cdots \otimes y_n) =
(-1)^{|x|} \sum_{i=1}^n x_{<i} \epsilon(y_{<i}) \otimes x_i \ast y_i \otimes \epsilon(x_{>i})y_{>i},
\end{align*}
where
\begin{align*}
x_{<i} & = x_1 \otimes \cdots \otimes x_{i-1}, &
y_{<i} & = y_1 \otimes \cdots \otimes y_{i-1}, \\
x_{>i} & = x_{i+1} \otimes \cdots \otimes x_n, & 
y_{>i} & = y_{i+1} \otimes \cdots \otimes y_n,
\end{align*}
with the convention
\begin{equation*}
x_{<1} = y_{<1} = x_{>n} = y_{>n} = 1 \in \Z,
\end{equation*}
and the only non-zero values of $x_i \ast y_i$ are
\begin{equation*}
\ast([0] \otimes [1]) = [0, 1], \qquad  \ast([1] \otimes [0]) = -[0, 1].
\end{equation*}

@@

\begin{theorem}
	The commutative diagram
	\begin{equation*}
	\begin{tikzcd}[column sep = small, row sep = small]
	&[18pt] U(\mathcal M) \arrow[dr, "\phi^\square", out=0] &[-0pt] \\
	\mathcal W \arrow[ur, "\psi_{U(\mathcal M)}", in=180] \arrow[rr, "\psi^\square"] & & \End_{N^\bullet (X)}
	\end{tikzcd}
	\end{equation*}
	defines a natural Steenrod-Adem structure on $N^\bullet(X)$ for any cubical set $X$.
\end{theorem}

>>>>

In this section we construct explicit May-Steenrod structures on three well know combinatorial $E_\infty$-operads: the Barratt-Eccles operad $\mathcal E$ \cite{berger04combinatorial}, the surjection operad $\mathcal X$ \cite{mcclure03cochain}, and the operad $U(\mathcal M)$ associated to the finitely presented $E_\infty$-PROP $\mathcal M$ \cite{medina2020prop1}. We also define natural and effective May-Steenrod structure on the normalized cochains of any simplicial or cubical set, by describing a $U(\mathcal M)$-algebra structures on it.

@@

\begin{figure}
	\begin{tikzcd}
	& & & U(\mathcal M) \arrow[dddr, bend left=10, "\phi^\triangle"'] \arrow[dddrr, bend left=20, "\phi^\square"'] & & \\
	& & \mathcal X \arrow[ru, "SL"] & & & \\
	& \mathcal E \arrow[ur, "TR"] & & & & \\
	\mathcal W \arrow[rrrr, "\psi^\triangle", bend right=10] \arrow[rrrrr, "\psi^\square", bend right=20] \arrow[ur, "\psi_{\mathcal E}"] \arrow[uurr, "\psi_{\mathcal X}"', bend right=35, pos=.6] \arrow[uuurrr, "\psi_{U(\mathcal M)}"', bend right=45, pos=0.6] & & & & \End_{N^\bullet(X)} & \End_{N^\bullet(X)}
	\end{tikzcd}
	\caption{Summary of effective constructions: May-Steenrod structures on the Barratt-Eccles $\mathcal E$, surjection $\mathcal X$, and $U(\mathcal M)$ operads, and natural May-Steenrod structures on the normalized chains of simplicial and cubical sets. We are abusing notation since the maps $TR$ and $SL$ required different sign conventions.}
	\label{fig: bigsummary}
\end{figure}

\subsection{Barratt-Eccles operad} In this subsection we effectively describe a May-Steenrod structure on the Barratt-Eccles operad.

@@

\begin{theorem} \label{thm: Steenrod-Adem on Barratt-Eccles}
	The morphism of $\mathrm{C}$-modules
	\begin{equation*}
	\psi_{\mathcal E} \colon \mathcal W \to \mathcal E
	\end{equation*}
	defines a May-Steenrod structure on the Barratt-Eccles operad.
\end{theorem}

@@

In this subsection we introduce a May-Steenrod structures on the surjection operad. There are two widely used sign conventions for this operad respectively due to McClure-Smith \cite{mcclure03cochain} and Berger-Fresse \cite{berger04combinatorial}. Using the Berger-Fresse convention, we can define a May-Steenrod structure on the surjection operad by composing the map $\psi_{\mathcal E}$ with the \textit{table reduction} quasi-isomorphism $\mathcal E \to \mathcal X$ introduced in Section 1.3 of \cite{berger04combinatorial}. We define an May-Steenrod structure on the surjection operad in a convention independent way which recovers the table reduction May-Steenrod structure in the Berger-Fresse case.

@@

\begin{theorem} \label{thm: Steenrod-Adem on surjection MS convention}
	The morphism of $\mathrm{C}$-modules
	\begin{equation*}
	\psi_{\mathcal X} \colon \mathcal W \to \mathcal X
	\end{equation*}
	defines a May-Steenrod structure on the Surjection operad.
\end{theorem}

@@

We will define a May-Steenrod structure $\psi_{U(\mathcal M)} \colon \mathcal W \to U(\mathcal M)$ by composing the May-Steenrod structure $\psi_{\mathcal X} : \mathcal W \to \mathcal X$ with a morphism of $\mathrm{C}$-modules ($\mathrm{S}$-modules in fact)
\begin{equation*}
SL \colon \mathcal X \to U(\mathcal M)
\end{equation*}
that we now define. Given a surjection $s \colon \{1, \dots, n+r\} \to \{1, \dots, r\}$ let $SL(s)$ be the element
\begin{center}
	\begin{tikzpicture}[scale=.85]
	\draw (0,0)--(0,-.6) node[below]{$\scriptstyle 1$};
	\draw (0,0)--(.5,.5);
	\draw (-.3, .3)-- (-.2,.5) node[above]{\quad $\scriptstyle 1\ 2\, ...\, k_1$};
	\draw (-.5,.5)--(0,0);
	\node at (.11,.4){$...$};

@@

\begin{theorem} \label{thm: Steenrod-Adem on U(M)}
	The composition $\psi_{U(\mathcal M)}$
	\begin{equation*}
	\begin{tikzcd}[column sep = normal]
	\mathcal W \arrow[r, "\psi_{\mathcal X}"] & \mathcal X \arrow[r, "SL"] & U(\mathcal M)
	\end{tikzcd}
	\end{equation*}
	defines a May-Steenrod structure on $U(\mathcal M)$.
\end{theorem}

@@

In this subsection we introduce a natural May-Steenrod structure on the normalized cochains of any simplicial set $X$. Since a May-Steenrod structure was constructed in the previous section for $U(\mathcal M)$, we only need to describe a natural $U(\mathcal M)$-algebra structure on $N^\bullet(X)$. Using the linear duality functor, it suffices to construct a natural $U(\mathcal M)$-coalgebra structure on $N_\bullet(X)$ which, in turn, can be derived via a Kan extension argument from one on each $N_\bullet(\triangle^n)$. We obtain these coalgebra structures by restricting a full $\mathcal M$-bialgebra structure. An $\mathcal M$-bialgebra structure is specified by three linear maps, the images of the generators
\begin{equation*}
\counit, \quad \coproduct, \quad \product,
\end{equation*}
satisfying the relations in the presentation of $\mathcal M$. For $n \in \mathbb{N}$, define: \vspace*{5pt} \\
(1) The counit $\epsilon \in \Hom(N_\bullet(\triangle^n), \Z)$ by
\begin{equation*}
\epsilon \big( [v_0, \dots, v_q] \big) = \begin{cases} 1 & \text{ if } q = 0, \\ 0 & \text{ if } q>0. \end{cases}
\end{equation*}
(2) The coproduct $\Delta \in \Hom(N_\bullet(\triangle^n), N_\bullet(\triangle^n)^{\otimes2})$ by
\begin{equation*}
\Delta \big( [v_0, \dots, v_q] \big) = \sum_{i=0}^q [v_0, \dots, v_i] \otimes [v_i, \dots, v_q].
\end{equation*}
(3) The product $\ast \in \Hom(N_\bullet(\triangle^n)^{\otimes 2}, N_\bullet(\triangle^n))$ by
\begin{equation*}
\left[v_0, \dots, v_p \right] \ast \left[v_{p+1}, \dots, v_q\right] = \begin{cases} (-1)^{p+|\pi|} \left[v_{\pi(0)}, \dots, v_{\pi(q)}\right] & \text{ if } v_i \neq v_j \text{ for } i \neq j, \\
0 & \text{ if not}, \end{cases}
\end{equation*}
where $\pi$ is the permutation that orders the totally ordered set of vertices, and $(-1)^{|\pi|}$ its sign.

@@

\begin{theorem}
	The commutative diagram
	\begin{equation*}
	\begin{tikzcd}[column sep = small, row sep = small]
	&[18pt] U(\mathcal M) \arrow[dr, "\phi^\triangle", out=0] &[-0pt] \\
	\mathcal W \arrow[ur, "\psi_{U(\mathcal M)}", in=180] \arrow[rr, "\psi^\triangle"] & & \mathrm{End}_{N^\bullet (X)}
	\end{tikzcd}
	\end{equation*}
	defines a natural May-Steenrod structure on $N^\bullet(X)$ for any simplicial set $X$.
\end{theorem}

@@

In this subsection we introduce a natural May-Steenrod structure on the normalized cochains of any cubical set. It follows the structure of the previous subsection closely. By the same considerations, the desired construction will follow from a natural $\mathcal M$-bialgebra structure on $N_\bullet(\square^n)$. These are determined by three linear maps satisfying the relations in the presentation of $\mathcal M$. For $n \in \mathbb{N}$, define: \vspace*{5pt} \\
(1) The counit $\epsilon \in \Hom(N_\bullet(\square^n), \Z)$ by
\begin{equation*}
\epsilon \left( x_1 \otimes \cdots \otimes x_d \right) = \epsilon(x_1) \cdots \, \epsilon(x_n),
\end{equation*}
where
\begin{equation*}
\epsilon([0]) = \epsilon([1]) = 1, \qquad \epsilon([0, 1]) = 0.
\end{equation*} \vspace*{-6pt} \\
(2) The coproduct $\Delta \in \Hom \left( N_\bullet(\square^n), N_\bullet(\square^n)^{\otimes 2} \right)$ by
\begin{equation*}	
\Delta (x_1 \otimes \cdots \otimes x_n) = 	
\sum \pm \left( x_1^{(1)} \otimes \cdots \otimes x_n^{(1)} \right) \otimes 	
\left( x_1^{(2)} \otimes \cdots \otimes x_n^{(2)} \right),	
\end{equation*}	
where the sign is determined using the Koszul convention, and we are using Sweedler's notation
\begin{equation*}	
\Delta(x_i) = \sum x_i^{(1)} \otimes x_i^{(2)}
\end{equation*}
for the chain map $\Delta \colon N_\bullet(\square^1) \to N_\bullet(\square^1)^{\otimes 2}$ defined by
\begin{equation*}
\Delta([0]) = [0] \otimes [0], \quad \Delta([1]) = [1] \otimes [1], \quad \Delta([0, 1]) = [0] \otimes [0, 1] + [0, 1] \otimes [1].
\end{equation*}
Using that $N_\bullet(\square^n) = N_\bullet(\square^1)^{\otimes n}$, $\Delta$ is the composition
\begin{equation*}
\begin{tikzcd}
N_\bullet(\square^1)^{\otimes n} \arrow[r, "\Delta^{\otimes n}"] & \left( N_\bullet(\square^1)^{\otimes 2}  \right)^{\otimes n} \arrow[r, "sh"] & \left( N_\bullet(\square^1)^{\otimes n} \right)^{\otimes 2}
\end{tikzcd}
\end{equation*}
where $sh$ is the shuffle map that places tensor factors in odd position first. \vspace*{5pt} \\
(3) The product $\ast \in \Hom(N_\bullet(\square^n)^{\otimes 2}, N_\bullet(\square^n))$ by
\begin{align*}
(x_1 \otimes \cdots \otimes x_n) \ast (y_1 \otimes \cdots \otimes y_n) =
(-1)^{|x|} \sum_{i=1}^n x_{<i} \epsilon(y_{<i}) \otimes x_i \ast y_i \otimes \epsilon(x_{>i})y_{>i},
\end{align*}
where
\begin{align*}
x_{<i} & = x_1 \otimes \cdots \otimes x_{i-1}, &
y_{<i} & = y_1 \otimes \cdots \otimes y_{i-1}, \\
x_{>i} & = x_{i+1} \otimes \cdots \otimes x_n, & 
y_{>i} & = y_{i+1} \otimes \cdots \otimes y_n,
\end{align*}
with the convention
\begin{equation*}
x_{<1} = y_{<1} = x_{>n} = y_{>n} = 1 \in \Z,
\end{equation*}
and the only non-zero values of $x_i \ast y_i$ are
\begin{equation*}
\ast([0] \otimes [1]) = [0, 1], \qquad  \ast([1] \otimes [0]) = -[0, 1].
\end{equation*}

@@

\begin{theorem}
	The commutative diagram
	\begin{equation*}
	\begin{tikzcd}[column sep = small, row sep = small]
	&[18pt] U(\mathcal M) \arrow[dr, "\phi^\square", out=0] &[-0pt] \\
	\mathcal W \arrow[ur, "\psi_{U(\mathcal M)}", in=180] \arrow[rr, "\psi^\square"] & & \End_{N^\bullet (X)}
	\end{tikzcd}
	\end{equation*}
	defines a natural May-Steenrod structure on $N^\bullet(X)$ for any cubical set $X$.
\end{theorem}

### 2021-05-25 introduction.tex
Rename to May-Steenrod operations

An outline of the article is described next.
We first introduce, in Section~2, the conventions we will follow regarding chain complexes, simplicial sets and cubical sets.
Then, in Section~3, we review the key notions from group homology we will use for cyclic and symmetric groups. Section~4 is devoted to the language of operads and related structures that we use, in Section~5, to define an integral structure at the chain level inducing Steenrod operations for every prime. Section~6, the bulk of this work, presents effective constructions of Steenrod-Adem structures on the Barratt-Eccles, surjection, and $U(\mathcal M)$ operads.
It also describes a natural $U(\mathcal M)$-algebra structure on the cochains of simplicial and cubical sets inducing natural Steenrod-Adem structures on them.
We end, in Section~7, with an overview of some connections of this work to important geometric and combinatorial structures, and provide an outline of future research directions.

>>>>

An outline of the article is described next.
We first introduce, in Section~2, the conventions we will follow regarding chain complexes, simplicial sets and cubical sets.
Then, in Section~3, we review the key notions from group homology we will use for cyclic and symmetric groups. Section~4 is devoted to the language of operads and related structures that we use, in Section~5, to define an integral structure at the chain level inducing Steenrod operations for every prime. Section~6, the bulk of this work, presents effective constructions of May-Steenrod structures on the Barratt-Eccles, surjection, and $U(\mathcal M)$ operads.
It also describes a natural $U(\mathcal M)$-algebra structure on the cochains of simplicial and cubical sets inducing natural May-Steenrod structures on them.
We end, in Section~7, with an overview of some connections of this work to important geometric and combinatorial structures, and provide an outline of future research directions.

### 2021-05-25 outlook.tex
Rename to May-Steenrod operations

This article looked at Steenrod operations from an algebraic viewpoint, a subject with rich geometric and combinatorial components as well. For example, \cite{Postnikov} and \cite{medina2018prop2} independently introduced equivalent geometric representations of the cup-$i$ products in terms of stabilized arc surfaces \cite{KLP} and weighted ribbon graphs respectively. In fact, an entire $E_\infty$-operad (PROP) is constructed geometrically in this way. We can also interpret the Steenrod-Adem structure in $U(\mathcal M)$ from the ``oriented surface" perspective, where the norm map $N$ of cyclic groups ---a key ingredient in Definition \ref{def: minimal cyclic resolution}--- was identified in \cite{KLP} with a Dehn twist operator in connection with Connes' cyclic complex.

>>>>

This article looked at Steenrod operations from an algebraic viewpoint, a subject with rich geometric and combinatorial components as well. For example, \cite{Postnikov} and \cite{medina2018prop2} independently introduced equivalent geometric representations of the cup-$i$ products in terms of stabilized arc surfaces \cite{KLP} and weighted ribbon graphs respectively. In fact, an entire $E_\infty$-operad (PROP) is constructed geometrically in this way. We can also interpret the May-Steenrod structure in $U(\mathcal M)$ from the ``oriented surface" perspective, where the norm map $N$ of cyclic groups ---a key ingredient in Definition \ref{def: minimal cyclic resolution}--- was identified in \cite{KLP} with a Dehn twist operator in connection with Connes' cyclic complex.
